### 2019-06-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
CHAUD

\paragraph{Two ideas for orthogonalization.}
\label{par:5:twoIdeasOrthogonalization}
%
On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
%
On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
% (for $k\in[K]$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
and selects at time $t$,
\begin{equation}
  A^j(t) := \pi\left(R^j(t), [g^j_\ell(t)]_{\ell=1,\dots,K}\right).
\end{equation}
If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.

% Now, our algorithms !!
\paragraph{Our two proposals.}
We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
\begin{equation}
  \TopM(t) := \left\{ \pi\left(k, \{g^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
\end{equation}

@@

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T - 1$}{
                  %
                  \eIf{$A^j(t) \notin \TopM(t)$}{
                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                      \tcp*[f]{randomly switch}
                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
                      }
                    }{
                      $A^j(t+1) = A^j(t)$
                      \tcp*[f]{stays on the same arm}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $g^j$).}
          \label{algo:5:RandTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

@@

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T-1$}{
                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
                      $A^j(t) \notin \TopM(t)$}
                    {
                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq g^j_{A^j(t)}(t-1)\right\}\right)$
                      \tcp*[f]{not empty} \\
                      % \tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}
                      $s^j(t+1) = \mathrm{False}$
                      \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
                    }
                    \uElseIf(\tcp*[f]{collision and not fixed}){
                        $C^j(t)$ \emph{and} $\overline{s^j(t)}$}
                      {
                        $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                        \tcp*[f]{transition $(2)$} \\
                        $s^j(t+1) = \mathrm{False}$
                    }
                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
                      $A^j(t+1) = A^j(t)$ \tcp*[f]{stay on the previous arm} \\
                      $s^j(t+1) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $g^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $g^j$).}
          \label{algo:5:MCTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

More precisely, under \MCTopM,
% \footnote{This choice is similar to the elementary step used in the \MusicalChair{} algorithm introduced by \cite{Rosenski16}, that gave its name to \MCTopM, even if they restrict to using the empirical averages as indices (\ie, the $0$-greedy algorithm).}
if player $j$ did not encounter a collision when using arm $k$ at time $t$,
then she marks her current arm as a ``chair'' ($s^j(t+1)=\mathrm{True}$),
and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
%
As soon as this ``chair'' $k$ is no longer in $\widehat{M_j}(t)$,
a new arm is sampled uniformly from a subset of $\TopM(t)$,
defined with the previous indices $g^j(t-1)$ (Lines~$3$-$5$).
%
The subset enforces a certain inequality on indices,
$g_{k'}^j(t-1) \leq g^j_{k}(t-1)$ and $g_{k'}^j(t) \geq g^j_{k}(t)$,
when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
This helps to control the number of such changes of arm,
as shown in Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}.
% the Appendix~\ref{proof:5:collisionsMCTopM}
The considered subset is never empty as it contains
at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
Collisions are dealt with only for non-fixed player $j$,
and when the previous arm is still in $\TopM(t)$.
%
In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
%
%
This stationary aspect helps to minimize the number of collisions,
as well as the number of switches of arm.
%
The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.

@@

  Thanks to \eqref{eq:5:KeyInclusion} and if $\Pr_{\bm{\mu}} $ denote here the probability under model $\bm{\mu}$, it is easy to decompose the number of selections of arm $k$ by user $j$ up to round $T$ as
  \begin{align*}
  \E_{\mu}[T_k^j(T)]
  &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
  = \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k \right).\\
  %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; g_{m^*}^j(t) < g_k^j(t) \right).
  %
  \end{align*}

  Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition

  \begin{align}
    \E_{\mu}[T_k^j(T)] &\leq
    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; g_{m_1^*}(t) < \mu_m^* \right) + \notag \\
    \;\;\;\;
    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; g_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  g_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(g_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
    %
    & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( g_{m_1^*}(t) < \mu_{m_1}^* \right)
    + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
    \label{eq:5:sumOfProbaForTkj}
  \end{align}
  where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in [M]$.

@@

  \begin{align*}
    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
  \end{align*}

  Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
  The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
  Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
  and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has

  \begin{align}
    \sum_{t=1}^T & \Pr\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
    & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
    & \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
    & \leq \sum_{s=1}^T \Pr\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
  \end{align}
  where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
  From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
  \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

  If $\mu_k > \mu_{k'}$, we rather use that
  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
  and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
  \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
  Summing over $k'$ yields the result.
\end{smallproof}

>>>>

\paragraph{Two ideas for orthogonalization.}
\label{par:5:twoIdeasOrthogonalization}
%
On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
%
On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
% (for $k\in[K]$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
and selects at time $t$,
\begin{equation}
  A^j(t) := \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right).
\end{equation}
If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.

% Now, our algorithms !!
\paragraph{Our two proposals.}
We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
\begin{equation}
  \TopM(t) := \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
\end{equation}

@@

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T - 1$}{
                  %
                  \eIf{$A^j(t) \notin \TopM(t)$}{
                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                      \tcp*[f]{randomly switch}
                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}){
                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      }
                    }{
                      $A^j(t+1) = A^j(t)$
                      \tcp*[f]{stays on the same arm}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:RandTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

@@

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T-1$}{
                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
                      $A^j(t) \notin \TopM(t)$}
                    {
                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : g_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      \tcp*[f]{not empty} \\
                      % \tcp*[f]{randomly switch on an arm that had smaller UCB at $t-1$}
                      $s^j(t+1) = \mathrm{False}$
                      \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
                    }
                    \uElseIf(\tcp*[f]{collision and not fixed}){
                        $C^j(t)$ \emph{and} $\overline{s^j(t)}$}
                      {
                        $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                        \tcp*[f]{transition $(2)$} \\
                        $s^j(t+1) = \mathrm{False}$
                    }
                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
                      $A^j(t+1) = A^j(t)$ \tcp*[f]{stay on the previous arm} \\
                      $s^j(t+1) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:MCTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

More precisely, under \MCTopM,
% \footnote{This choice is similar to the elementary step used in the \MusicalChair{} algorithm introduced by \cite{Rosenski16}, that gave its name to \MCTopM, even if they restrict to using the empirical averages as indices (\ie, the $0$-greedy algorithm).}
if player $j$ did not encounter a collision when using arm $k$ at time $t$,
then she marks her current arm as a ``chair'' ($s^j(t+1)=\mathrm{True}$),
and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
%
As soon as this ``chair'' $k$ is no longer in $\widehat{M_j}(t)$,
a new arm is sampled uniformly from a subset of $\TopM(t)$,
defined with the previous indices $U^j(t-1)$ (Lines~$3$-$5$).
%
The subset enforces a certain inequality on indices,
$U_{k'}^j(t-1) \leq U^j_{k}(t-1)$ and $U_{k'}^j(t) \geq U^j_{k}(t)$,
when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
This helps to control the number of such changes of arm,
as shown in Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}.
% the Appendix~\ref{proof:5:collisionsMCTopM}
The considered subset is never empty as it contains
at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
Collisions are dealt with only for non-fixed player $j$,
and when the previous arm is still in $\TopM(t)$.
%
In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
%
%
This stationary aspect helps to minimize the number of collisions,
as well as the number of switches of arm.
%
The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.

@@

  Thanks to \eqref{eq:5:KeyInclusion} and if $\Pr_{\bm{\mu}} $ denote here the probability under model $\bm{\mu}$, it is easy to decompose the number of selections of arm $k$ by user $j$ up to round $T$ as
  \begin{align*}
  \E_{\mu}[T_k^j(T)]
  &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
  = \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k \right).\\
  %&= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
  &= \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\;\; \exists m \in[M]:\; U_{m^*}^j(t) < g_k^j(t) \right).
  %
  \end{align*}

  Considering the relative position of the upper-confidence bound $U_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition

  \begin{align}
    \E_{\mu}[T_k^j(T)] &\leq
    \sum_{t=1}^T\Pr_{\bm{\mu}} \left( \exists m_1 \in [M]: \; U_{m_1^*}(t) < \mu_m^* \right) + \notag \\
    \;\;\;\;
    \sum_{t=1}^T\Pr_{\bm{\mu}} & \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; U_{m_2^*}(t) \leq g_k(t) , \forall {m_3} \in [M]: \;  U_{{m_3}^*}(t) \geq \mu_{m_3}^* \right) \notag \\
    &\leq \sum_{m_1=1}^M\sum_{t=1}^T\Pr_{\bm{\mu}} \left(U_{m_1^*}(t) < \mu_{m_1}^* \right) + \sum_{t=1}^T\Pr_{\bm{\mu}} \left(A^j(t) = k,\;\; \exists m_2 \in [M]:\; \mu_{m_2}^* \leq g_k(t)\right) \notag \\
    %
    & \leq \sum_{m_1=1}^M \sum_{t=1}^T \Pr_{\bm{\mu}} \left( U_{m_1^*}(t) < \mu_{m_1}^* \right)
    + \sum_{t=1}^T \Pr_{\bm{\mu}} \left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
    \label{eq:5:sumOfProbaForTkj}
  \end{align}
  where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in [M]$.

@@

  \begin{align*}
    & \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \Pr\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': U_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
    & \leq   \sum_{t=1}^T \sum_{k' \neq k}\Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right) \\
    & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \Pr\left(A^j(t) = k,A^j(t+1) = k', U_{k'}^j(t) \geq U_{k}^j(t), U_{k'}^j(t-1) \leq U_{k}^j(t-1)  \right)}_{:= T_{k'}}
  \end{align*}

  Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right)\]
  The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
  Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
  and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has

  \begin{align}
    \sum_{t=1}^T & \Pr\left(A^j(t) = k, U_{k}^j(t-1) \geq \mu_{k'}\right) \nonumber\\
    & =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
    & \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
    & \leq \sum_{s=1}^T \Pr\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
  \end{align}
  where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
  From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
  \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

  If $\mu_k > \mu_{k'}$, we rather use that
  \[T_{k'} \leq \sum_{t=1}^T \Pr\left(U_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \Pr\left(A^j(t+1) = k', U_{k'}^j(t) \geq \mu_{k}\right)\]
  and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
  \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
  Summing over $k'$ yields the result.
\end{smallproof}
